---
title: "How to set up a package for continuous integration"
output:
  html_document:
    df_print: paged
---

# R package basics

First, create a new package, e.g. using the RStudio GUI. The *usethis* package provides all upcoming steps immensely:

```{r eval=FALSE}
library(usethis)
```

Feel free to delete the hello world example (R and Rd file). We'll add the functionality via

```{r eval=FALSE}
use_r("add.R")
```

The code we add is as follows:

```{r eval=FALSE}
add <- function(a,b) {
  return(a+b)
}
```


To see the new function in action run "Install and Restart" from the Build tab:

```{r eval=FALSE}
add(7,2)
```

We'll now add more functionality:

```{r eval=FALSE}
add <- function(a,b,plot=FALSE) {
  val <- a+b
  if (plot==TRUE) {
    df <- data.frame(x=c(a,b,val))
    p <- ggplot2::ggplot(df) + ggplot2::geom_col(ggplot2::aes(x=x,y=x))
    print(p)
  }
  return(val)
}
```


Use "Source on save" to play around with the function. When done, install and restart again.

```{r eval=FALSE}
add(4,16,plot=TRUE)
```


The package now depends on ggplot2. We have to add this to the description file

```{r eval=FALSE}
use_package("ggplot2", "Imports")
```

As usethis tells you to do, always refer to functions with `ggplot2::fun()`.

Finally, we'll ad a license to our package

```{r eval=FALSE}
use_gpl3_license(name = "Oliver Pfaffel") # required to share improvements
```

If you are a pipe fan, you can easily make it available via

```{r eval=FALSE}
use_pipe() # Use %>%
```



## Documentation

We'll add a news and readme file. 

```{r eval=FALSE}
use_news_md()
use_readme_md() # use_readme_rmd() IF you want to run R code in your readme
```

We change the readme file as follows.

```{r eval=FALSE}
# CIexamplePkg

<!-- badges: start -->
<!-- badges: end -->

The goal of CIexamplePkg is to provide a simple example on how to set up a package for continuous integration.
```

Be sure to keep the news file up to date. You can increment the version number via

```{r eval=FALSE}
use_version()
```

You can also add a vignette, but we'll not do this here.

```{r eval=FALSE}
use_vignette("name_of_vignette")
```

Every R function should have a documentation. We'll change the code of our function to

```{r eval=FALSE}
#' Sum up two variables
#'
#' What is the sum of a and b?
#'
#' @param a numeric
#' @param b numeric
#' @param plot Makes plot if TRUE
#'
#' @return Returns a numeric that is the sum of a and b.
#'
#' @examples
#' add(7,2)
#' add(4,16,plot=TRUE)
#'
#' @export
add <- function(a,b,plot=FALSE) {
  val <- a+b
  if (plot==TRUE) {
    df <- data.frame(x=c(a,b,val))
    p <- ggplot2::ggplot(df) + ggplot2::geom_col(ggplot2::aes(x=x,y=x))
    print(p)
  }
  return(val)
}
```


The function will now be exported to the namespace of our package, if we delete the existing NAMESPACE file first.
After "Install and Restart" we build the documentation via

```{r eval=FALSE}
devtools::document()
```

Now the documentation is available via

```{r eval=FALSE}
?add
```


## Checking the package

Now we should check the package (there also is a button for this in RStudio)

```{r eval=FALSE}
devtools::check(document = FALSE)
```

To remove the note about non-standard files on top level, we simply ignore those file when building R

```{r eval=FALSE}
use_build_ignore(c("setup_CI.Rmd'","setup_CI.nb.html"))
```

## Tests

Every R package should have tests that automatically check core functionality

```{r eval=FALSE}
use_test("add")
```

We'll simply change the code to

```{r eval=FALSE}
test_that("addition works works", {
  expect_equal(2 + 2, 4)
})
```

and run the test. It should be successfull.

# Continous integration

We'll make use of the travis package here

```{r eval=FALSE}
remotes::install_github("ropenscilabs/travis")
```

## git(hub)

```{r}
usethis::use_git()
usethis::use_git_config(user.name = "Oliver Pfaffel", user.email = "opfaffel@gmail.com")
```

Now add a github PAT to 



alternatively use travis package
Create a repository on GitHub (if it's not there yet)
travis::github_create_repo()

Show the GitHub repository name
travis::github_repo()


We'll make use of the travis package here

```{r eval=FALSE}
devtools::install_github("ropenscilabs/tic.package")
```


```{r}
tic.package::
```

